VM-Centric Deduplication with Fault Isolation for Snapshot Backup in the Cloud

### 2014-01-29 snapshot_system/paper_analysis.tex
revise

\begin{table}[htbp]
\centering
\tabcolsep=0.11cm
\begin{tabular}{|p{1.0cm}|p{5.75cm}|}
%\hline
%$k$ &  the number of top most popular chunks selected for deduplication\\ 
%\hline
%$c$ &  the total amount of data chunks in a cluster of VMs\\ 
%\hline
%$c_u$ &  the total amount of unique fingerprints after perfect  deduplication\\
%\hline
%$f_i$ &  the frequency for the $i$th most popular fingerprint\\
%\hline
%$\delta$ &  the percentage of duplicates detected in local deduplication\\
%\hline
%$\sigma$ & = percentage of unique data  belonging to  PDS\\
\hline
$r$, $r_c$ & replication degree of non-PDS, and PDS file blocks\\
\hline
$n$, $p$ & no. of virtual and physical machines in the cluster\\
%\hline
%$V$ & the average number of VMs per machine\\
%\hline
%$E_c, E_o$ & deduplication efficiency of VC and VO \\
%\hline
%$D$ & the amount of unique data on each machine\\
%\hline
%$s$ & the average data chunk size. Our setting is  4K.\\
%\hline
%$s$ & the average number of chunks per FSB\\
%\hline
%$m$ & memory size on each node used by VC\\ 
%\hline
%$E$ & the size of an popular data index entry\\
\hline
$N_1$, $N_2$ & the average number  of non-PDS and PDS file blocks in a VM for VC\\
\hline
$N_o$ & the average number  of file blocks  in a VM for VO\\
\hline
A(r) & availability of a file block with $r$ replicas and $d$ failed physical machines\\
\hline
\end{tabular}
\caption{Modeling  parameters}
\label{tab:symbol}
\end{table}

@@

For VC, each VM has an average number $N_1$ of non-PDS FSBs
and  has  an average of  $N_2$ PDS FSBs. 
Each non-PDS FSB is associated with one VM
 and  we denote that PDS FSBs are
shared by an average of $V_c$ VMs. 
Then, 
\[
n* N_1   \approx \mbox{ No of non-PDS file system blocks} 
\]

For VO, each VM has an average of $N_o$ file system blocks
and let $V_o$ be the average number of VMs shared by each FSB.
\[
n *N_o   =  V_o *\mbox{ No of file system blocks} 
\]

@@

\begin{figure}[htbp]
  \centering
	\begin{tikzpicture}
		\begin{axis}[
		%title={VO VM links},
                width=\linewidth,
                height=0.6\linewidth,
		xlabel={Number of VMs},
		ylabel={Avg. Num. FSBs used by VM},
                xmin=0,
                ymin=0,
                xmax=106,
		%extra y ticks={4.5,5.5,6.5} %to show extra ticks
		%legend pos=north east,
                legend style={
                    at={(1,0.825)},
                    anchor=north east
                },
                %legend columns=3
		]
                \addplot[blue,mark=none] table[x=VMs,y=N_O] {figures/vm_links_all.txt};
                %\addplot[red,dotted,mark=none] table[x=VMs,y expr=\thisrow{N_1}+\thisrow{N_2}] {figures/vm_links_all.txt};
                \addplot[red,dashdotted] table[x=VMs,y=N_1] {figures/vm_links_all.txt};
                \addplot[red,densely dashed,mark=none] table[x=VMs,y=N_2] {figures/vm_links_all.txt};
                \legend{$N_o$,
                    %$N_1+N_2$,
                    $N_1$,
                    $N_2$};
		\end{axis}
	\end{tikzpicture}
  \caption{Measured average number of 64MB FSBs used by a single VM. For VC both the number of PDS and Non-PDS FSBs used are shown.}
  \label{fig:vm-links}
\end{figure}

>>>>

\begin{table}[htbp]
\centering
\tabcolsep=0.11cm
\begin{tabular}{|p{1.0cm}|p{5.75cm}|}
%\hline
%$k$ &  the number of top most popular chunks selected for deduplication\\ 
%\hline
%$c$ &  the total amount of data chunks in a cluster of VMs\\ 
%\hline
%$c_u$ &  the total amount of unique fingerprints after perfect  deduplication\\
%\hline
%$f_i$ &  the frequency for the $i$th most popular fingerprint\\
%\hline
%$\delta$ &  the percentage of duplicates detected in local deduplication\\
%\hline
%$\sigma$ & = percentage of unique data  belonging to  PDS\\
\hline
$r$, $r_c$ & replication degree of non-PDS, and PDS file blocks in VC\\
\hline
$n$, $p$ & no. of virtual and physical machines in the cluster\\
%\hline
%$V$ & the average number of VMs per machine\\
%\hline
%$E_c, E_o$ & deduplication efficiency of VC and VO \\
%\hline
%$D$ & the amount of unique data on each machine\\
%\hline
%$s$ & the average data chunk size. Our setting is  4K.\\
%\hline
%$s$ & the average number of chunks per FSB\\
%\hline
%$m$ & memory size on each node used by VC\\ 
%\hline
%$E$ & the size of an popular data index entry\\
\hline
$N_1$, $N_2$ & the average number  of non-PDS and PDS file blocks in a VM for VC\\
\hline
$N_o$ & the average number  of file blocks  in a VM for VO\\
\hline
A(r) & availability of a file block with $r$ replicas and $d$ failed physical machines\\
\hline
\end{tabular}
\caption{Modeling  parameters}
\label{tab:symbol}
\end{table}

@@

For VC, each VM has an average number $N_1$ of non-PDS file system blocks
and  has  an average of  $N_2$ PDS file system blocks. 
Each non-PDS block is associated with one VM
 and  we denote that PDS blocks are
shared by an average of $V_c$ VMs. 
Then, 
\[
n* N_1   \approx \mbox{ No of non-PDS file system blocks} 
\]

For VO, each VM has an average of $N_o$ file system blocks
and let $V_o$ be the average number of VMs shared by each file block.
\[
n *N_o   =  V_o *\mbox{ No of file system blocks} 
\]

@@

\begin{figure}[htbp]
  \centering
	\begin{tikzpicture}
		\begin{axis}[
		%title={VO VM links},
                width=\linewidth,
                height=0.6\linewidth,
		xlabel={Number of VMs},
		ylabel={Avg. no file system blocks  used by VM},
                xmin=0,
                ymin=0,
                xmax=106,
		%extra y ticks={4.5,5.5,6.5} %to show extra ticks
		%legend pos=north east,
                legend style={
                    at={(1,0.825)},
                    anchor=north east
                },
                %legend columns=3
		]
                \addplot[blue,mark=none] table[x=VMs,y=N_O] {figures/vm_links_all.txt};
                %\addplot[red,dotted,mark=none] table[x=VMs,y expr=\thisrow{N_1}+\thisrow{N_2}] {figures/vm_links_all.txt};
                \addplot[red,dashdotted] table[x=VMs,y=N_1] {figures/vm_links_all.txt};
                \addplot[red,densely dashed,mark=none] table[x=VMs,y=N_2] {figures/vm_links_all.txt};
                \legend{$N_o$,
                    %$N_1+N_2$,
                    $N_1$,
                    $N_2$};
		\end{axis}
	\end{tikzpicture}
  \caption{Measured average number of 64MB file system blocks used by a single VM in
VC and VO. }
  \label{fig:vm-links}
\end{figure}

### 2014-01-29 snapshot_system/paper_background.tex
revise

%minimize data sharing and simplify deletion while sacrificing 
%deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\paragraph*{Units of Storage versus Units of Sharing} --
The file system block (FSB) size in a distributed file system such as  
the Hadoop file system and the Google file system
is uniform and large (e.g.  64MB) so as to optimize file system performance
for bandwidth usage.
Typically, deduplication implementations use smaller block sizes (4K bytes to
8K bytes on average).
In addition to the failure correlation that cross-VM data deduplication can
introduce, disk block sharing of data chunks is another potential source of
hidden data dependence.
Thus a VM-centric solution should consider how data chunks from VMs are
distributed among large file system blocks when a large-scale distributed file
system is used for disk image files.

>>>>

%minimize data sharing and simplify deletion while sacrificing 
%deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\paragraph*{Units of Storage versus Units of Sharing} --
The file system block  size in a distributed file system such as  
the Hadoop file system and the Google file system~\cite{googlefs03}
is large (e.g.  64MB) so as to optimize file system performance
for big data.
Typically, deduplication implementations use smaller block sizes (4K bytes to
8K bytes on average~\cite{Jin2009}).
In addition to the failure correlation that cross-VM data deduplication can
introduce, disk block sharing of data chunks is another potential source of
hidden data dependence.
Thus a VM-centric solution should consider how data chunks from VMs are
distributed among large file system blocks when a large-scale distributed file
system is used for disk image files.

### 2014-01-29 snapshot_system/paper_intro.tex
revise

Commercial ``Infrastructure as a Service'' clouds (i.e. {\em public
clouds}) often make use of commodity data center components to achieve the
best possible economies of scale.  In particular, large-scale e-commerce cloud
providers such as Google and Alibaba deploy ``converged'' components that
co-locate computing and storage in each hardware module (as opposed to having
separate computing and storage ``tiers.'')  The advantage of such an approach
is that all infrastructure components are used to support paying customers --
there are no resources specifically dedicated to cloud services.
%One of emerging architectures for building cloud services is a converged
%storage architecture that leverages commodity servers with software clustered
%storage.
%The converged architecture (server plus compute in same tier) grew out of the
%Google model. 
In particular, these providers use software to
aggregate multiple direct attached low-cost disks together across
servers as a way of avoiding the relatively high cost of network attached 
storage~\cite{googlefs03,hdfs10}.
%systems. %This(e.g. ~\cite{NutanixPaper}).  Leading cloud platform companies
%such as Google, Amazon, Alibaba, Microsoft Azure  have built a converged
%compute and storage infrastructure and used a distributed file system such as
%Google file system~\cite{googlefs03,hdfs10} to glue a large number of
%commodity servers.
% with local storage into a single cluster.
In such an environment,
%That  allows the use of scalable compute and storage, without incurring the
%costs and performance limitations associated with network storage.
each physical machine runs a number of virtual machines (VMs)
%as  instances of a guest operating system 
and their virtual disks are stored as disk image files, typically
using a local (possibly shared) file system.
Frequent snapshot backup of virtual disk images can
increase the service reliability by allowing VMs to restart from their latest
snapshot in the event of a server failure.  However, because the disk
blocks that make up the VMs are often the same from snapshot to snapshot,  
deduplication of redundant content
blocks~\cite{venti02,bottleneck08} is necessary to substantially reduce the
storage demand.
%For example, the Aliyun cloud, which is  the largest cloud service provider
%by Alibaba in China, automatically conducts  the backup of virtual disk
%images to all active users every day.  The cost of supporting a large number
%of concurrent backup streams is high because of the huge storage demand and
%the use of deduplication

Without deduplication,
the cloud typically offloads backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out pre-deduplication backup data wastes huge amount
of network bandwidth that would otherwise be available to user VMs. 
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).

In this paper we propose a backup service for converged cloud architectures
that is co-located with user VMs, sharing the cloud computer, network, and
storage resource with them.
We describe methodologies to address the following challenges: 
\begin{itemize}
\item good global deduplication efficiency using a distributed implementation
that limits its resource usage, 
\item augmented fault isolation for the sharing of deduplicated data,
given that node failures in converged architecture are the norm rather than
exception, and
\item efficient garbage collection so that users stop incurring 
storage charges immediately after snapshot deletion.
\end{itemize}

@@

Version-based detection has been used to identify file blocks that have not
changed across snapshots so their redundant storage can be
avoided~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
%the previous version of the
%snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}, 
Alternatively, techniques that compare block ``fingerprints'' to
identify duplicates
that exist among all files regardless of their 
origin~\cite{Guo2011,Dong2011,extreme_binning09} have
also gained in popularity.

One consequence of aggressive deduplication is the possible loss of failure
resilience.  Separate files share the same physical copies of blocks that are
logically duplicated among them.  If a shared block is lost, all files that
share that block are affected.  The problem is particularly acute for VM
snapshots where the purpose of the snapshot is precisely to create redundant
data copies so that VMs are more fault resilient.  Thus, there exists a
tension between the additional redundancy that snapshot creation is intended
to provide, and the storage efficiencies that deduplication makes possible.

%Because
%of highly repetitive content in snapshots from different VMs, many data chunks
%are shared by virtual machines.  Failure of a few shared data chunks can have
%a broad effect and snapshots of these virtual machines could be affected.  The
%previous work in deduplication focuses on the efficiency and approximation of
%fingerprint comparison, and has not addressed fault tolerance issues  together
%with deduplication.

The main technical approach of our work focuses on
sharing only ``popular'' data blocks 
(through deduplication and limited replication)
across 
virtual machines while using localized deduplication within each VM
to achieve both storage savings and fault isolation.
%, and to propose a VM-centric approach that localizes
%deduplication as much as possible and restricts global deduplication only to a
%limited set of most popular blocks.  
%In our method, local deduplication uses
%similarity-guided elimination to improve the deduplication coverage.  
Since
the file system block size is normally bigger than the average data chunk size
used for deduplication, 
our approach uses a deduplication data chunk size that is smaller than a
single file block and
packages data chunks from the same VM into a file
system block as much as possible to improve fault isolation.  Deduplicating
data locally within a VM also reduces the need to maintain a global index of
chunks that are shared among VMs, thereby increasing the scaling efficiency of
the snapshot system.  The cost associated with localization, however, is in 
the lost
storage efficiency that cross-VM deduplication makes possible.  To recover
some of this efficiency we allow a small number of the most popular data
blocks to be deduplicated (and then replicated) across VMs.

This paper explores the storage and
performance efficiencies that such a {\em VM-centric} 
combination of local and global
techniques can achieve.  We term it VM-centric because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
We detail the storage, performance,
and fault isolation properties
of our method and verify these properties empirically using a prototype
system that we have developed to run as a co-located back-up service in
converged IaaS cloud environments.

>>>>

Commercial ``Infrastructure as a Service'' clouds (i.e. {\em public
clouds}) often make use of commodity data center components to achieve the
best possible economies of scale.  In particular, large-scale e-commerce cloud
providers such as Google and Alibaba deploy ``converged'' components that
co-locate computing and storage in each hardware module (as opposed to having
separate computing and storage ``tiers.'')  The advantage of such an approach
is that all infrastructure components are used to support paying customers --
there are no resources specifically dedicated to cloud services.
%One of emerging architectures for building cloud services is a converged
%storage architecture that leverages commodity servers with software clustered
%storage.
%The converged architecture (server plus compute in same tier) grew out of the
%Google model. 
In particular, these providers use software to
aggregate multiple direct attached low-cost disks together across
servers as a way of avoiding the relatively high cost of network attached 
storage~\cite{googlefs03,hdfs10}.
%systems. %This(e.g. ~\cite{NutanixPaper}).  Leading cloud platform companies
%such as Google, Amazon, Alibaba, Microsoft Azure  have built a converged
%compute and storage infrastructure and used a distributed file system such as
%Google file system~\cite{googlefs03,hdfs10} to glue a large number of
%commodity servers.
% with local storage into a single cluster.
In such an environment,
%That  allows the use of scalable compute and storage, without incurring the
%costs and performance limitations associated with network storage.
each physical machine runs a number of virtual machines (VMs)
%as  instances of a guest operating system 
and their virtual disks are stored as disk image files, typically
using a local (possibly shared) file system.
Frequent snapshot backup of virtual disk images can
increase the service reliability by allowing VMs to restart from their latest
snapshot in the event of a server failure.  However, because the disk
blocks that make up the VMs are often the same from snapshot to snapshot
and are shared among VMs,  
deduplication of redundant content
blocks~\cite{venti02,bottleneck08} is necessary to substantially reduce the
storage demand.
%For example, the Aliyun cloud, which is  the largest cloud service provider
%by Alibaba in China, automatically conducts  the backup of virtual disk
%images to all active users every day.  The cost of supporting a large number
%of concurrent backup streams is high because of the huge storage demand and
%the use of deduplication

%Without deduplication,
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out undeduplicated backup data wastes huge amount
of network bandwidth that would otherwise be available to user VMs. 
%Even when the storage system implements deduplication internally
%such dedicated backup storage is either difficult to scale out
%(which can comprise the overall scalability of the converged architecture) 
%or it must forego global deduplication in order to gain scalability 
%(which increases backup cost).

In this paper we consider a backup service for converged cloud architectures
that is co-located with user VMs, sharing the cloud computer, network, and
storage resource with them.
We describe methodologies to address the following two challenges: 
\begin{itemize}
\item simple deduplication and snapshot deletion with competitive efficiency and 
a small resource usage without affecting other collocated cloud services.
\item augmented fault isolation for the sharing of deduplicated data,
given that node failures in converged architecture are the norm rather than
exception. 
% so that users stop incurring storage charges immediately after snapshot deletion.
\end{itemize}

@@

Version-based detection has been used to identify file blocks that have not
changed across snapshots so their redundant storage can be
avoided~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
%the previous version of the
%snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}, 
Alternatively, techniques that compare block ``fingerprints'' to
identify duplicates
that exist among all files regardless of their 
origin~\cite{Guo2011,Dong2011,extreme_binning09} have
also gained in popularity.  
One consequence of aggressive deduplication is the possible loss of failure
resilience.  Separate files share the same physical copies of blocks that are
logically duplicated among them.  If a shared block is lost, all files that
share that block are affected.  
%The problem is particularly acute for VM
%snapshots where the purpose of the snapshot is precisely to create redundant
%data copies so that VMs are more fault resilient.  
Thus, there exists a
tension between the additional redundancy that snapshot creation is intended
to provide, and the storage efficiency that deduplication makes possible.
The previous work on deduplication has not considered the impact of 
duplicate sharing on  fault tolerance.
%Because
%of highly repetitive content in snapshots from different VMs, many data chunks
%are shared by virtual machines.  Failure of a few shared data chunks can have
%a broad effect and snapshots of these virtual machines could be affected.  The
%previous work in deduplication focuses on the efficiency and approximation of
%fingerprint comparison, and has not addressed fault tolerance issues  together
%with deduplication.

%This paper explores the storage and
%performance efficiencies that such a {\em VM-centric} 
%combination of local and global
%techniques can achieve.  
We term our approach {\em VM-centric} because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
We detail the storage, performance,
and fault isolation properties
of our method and verify these properties empirically using a prototype
system that we have developed to run as a co-located back-up service in
converged IaaS cloud environments.
Our work focuses on sharing only ``popular'' data blocks 
(through deduplication and limited replication)
across 
virtual machines while using localized deduplication within each VM
to achieve both storage savings and fault isolation.
%, and to propose a VM-centric approach that localizes
%deduplication as much as possible and restricts global deduplication only to a
%limited set of most popular blocks.  
%In our method, local deduplication uses
%similarity-guided elimination to improve the deduplication coverage.  
Since
the file system block size is normally bigger than the average data chunk size
used for deduplication,  packaging
content from different VMs into the same file block  creates a ``false'' sharing that 
affects fault isolation. Thus we avoid such VM-oblivious packaging.
%our approach uses a deduplication data chunk size that is smaller than a
%single file block and
%Thus our strategy is to packages data chunks from the same VM into a file
%system block as much as possible to improve fault isolation.  
%Deduplicating
%data locally within a VM also reduces the need to maintain a global index of
%chunks that are shared among VMs, thereby increasing the scaling efficiency of
%the snapshot system.  
The cost associated with localization, however, is in 
the lost
storage efficiency that cross-VM deduplication makes possible.  To recover
some of this efficiency we allow a small number of the most popular data
blocks to be deduplicated (and then replicated) across VMs.
